\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-06-05 trunk/arch.tex
updates

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.45]{figs/token-disptch.pdf}
  \caption{\small \kcomment{Dispatching a token occurs only when the
      request start time overlaps a window. \texttt{Goal$_1$}
      therefore gets dispatched from \texttt{Reactor$_1$} to
      \texttt{Reactor$_2$} but not \texttt{Goal$_2$} to avoid early
      commitment. Tokens from \texttt{Reactor$_1$} are 'dispatched'
      for execution to \texttt{Reactor$_2$} only when necessary while
      meeting the conditions in Definition \ref{def:dispatch}. $\tau$
      signifies the execution frontier as time moves from left to
      right in the illustration.}}
  \label{fig:reactor-lookahead}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines).]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

% \label{sec:delib-exec-as}
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \subfloat[\small \kcomment{A view of synchronization and deliberation within a
  tick cycle.}]{\label{fig:sync-cycle}\includegraphics[width=0.35\columnwidth]{figs/sync-tick-cycle}} 
  \qquad
  \subfloat[\small A \rx reactor's execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.]{\label{fig:tick-exec}\includegraphics[width=0.45\columnwidth]{figs/tick-cycle}}
    \caption{\small \kcomment{Detailed and abstracted views of
        synchronization and deliberation within one and multiple tick
        cycles.}}
  \label{fig:sync-plan-cycle}
\end{figure}

@@

\begin{figure}[b]
  \centering \subfloat[\small When the \texttt{Sampling} reactor fails
  to synchronize, it is removed from the reactor dependency graph
  along with all its {\em internal} timelines while generating a
  \texttt{Failed} observation.]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.pdf}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize too leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent \kcomment{from} Fig. \ref{fig:agent}.}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of synchronization flaws in a
    reactor. The reactor receives new observations when they are
    produced by the owner(s) of its internal timelines. The open
    ``holes'' after the last token of each timeline represents the
    domain of all possible values for the end of this token. At every
    tick $\tau$ the reactor needs to integrate the {\em External}
    state information it received and -- based on its model
    $\mathcal{M}$ -- resolve its {\em Internal} state that will then
    be provided by the agent to other reactors using these state
    variables.}
  \label{fig:synch:flaw}
\end{figure}

@@

Steps $2$ \& $3$ above are handled by the \eu solver from Algorithm
\ref{alg:europa:solve} which ensures that the {\em current state
  flaws} are selected last so as to follow the sequence given above.

As {\em synchronization} focuses on the execution frontier it also
filters out all flaws that are outside of the current tick's scope.
If the flaw correlates to a token that is {\em necessarily} ending
before $\tau$ or starting after $\tau+1$ then the solver ignores
it. Such a strategy helps to reduce the scope of synchronization
exclusively to the current state of the reactor and in turn reduces
the search space for the {\em synchronization} solver.  To do so we
need to assume that the current state value of an {\em internal}
timeline does not depend on the future or more accurately that any
choice made during this {\em synchronization} will not lead to a
future inconsistency.  Such an assumption impacts the set of possible
domains the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model:

\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   equals(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11 }
\end{verbatim}

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially since it
puts a strong emphasis between an evolving plan and an unknown future;
\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above. For instance, assume that the \texttt{Agent.location} is
{\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on the \texttt{Agent.location} timeline, is to be
\texttt{At(SuperMarket)} in the directly foreseeable future.

In the real world the fact that the robot is \texttt{Going} to a
certain location does not necessarily imply that it will succeed in
ending up there. Due to an exogenous event, for example, the road to
the \texttt{SuperMarket} could be closed, which could result in an
alternate observation that the robot is \texttt{At(7th st. \& Main)}
contradicting the existing plan. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes the
contradiction of the \texttt{At} predicate, the rule that generated
this conflict is connected to past observations \texttt{Going} and
\texttt{Go}. These cannot be retracted easily without a potential
impact to the plan implying in terms of a cost in back-propagation.

@@

One could therefore alter this part of the model to give the reactor
more flexibility as follows:

% \begin{minipage}[t]{\textwidth}
% \begin{figure}
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
% \label{lab:model:shopping}
% \caption{\small Model fragment to enforce a \texttt{Going} end condition}
% \end{figure}
% \end{minipage}

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants the robot to be at. This first level
of relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.

As the example above shows, synchronization involves a deliberation
process. The risk is such that the search for a solution is not a
simple sequence of decisions where the resolution can be easily
identified by the solver as any other option would lead immediately to
plan inconsistency. Since such a process is required by the \rx at
\emph{every} tick for \emph{every} reactor the model needs to be
designed carefully.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the \rx agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. This special state is a special predicate which is
acceptable for all the {\em internal} and {\em external} timelines of
\kcomment{an} \eu reactor.  By so doing we allow for graceful
degradation where other reactors can readapt their plans on receipt of
this observation.

Finally, at the engine level careful choices were made in both the
focus of the search -- limited only to tokens that can potentially
overlap the current tick -- and heuristics where the {\em current
  state flaw} is considered last during the search process. The later
limits the amount of backtracking for a solution -- or at least how
deep one needs to backtrack in a decision tree.  Similarly the
sequence in which we evaluate the possible resolution for such a flaw
-- presented in Proposition \ref{prop:csf:resolve} -- was selected
according to some simple assumptions, that any external observation
that is compatible with an existing token in the current plan reflects
the execution of the plan itself. In other words this means that our
solver always prefers to use tokens \emph{already existing} in the
plan as a way to identify current internal state. By doing so
synchronization exploits the structure of the plan produced during
previous deliberation steps.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window}
  \caption{\small The planning window of the deliberation solver and
    its relation to the {\em dispatch window}. The dark grey tokens
    with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$. All the other
    tokens ({\color{green} green}) along with the goals received from
    the agent ({\color{blue} blue}) which can overlap this window, are
    evaluated.}
  \label{fig:plan:window}
\end{figure}

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch window} from
Definition \ref{def:dispatch} which filters out at the agent level
those goals that are before the reactor execution latency or too far
into the future.  The planning window refines this filtering further,
as tokens can be excluded dynamically from deliberation as their
insertion in the plan and scheduling with other generated tokens can
alter their temporal scope to be outside of the planning window.  This
allows the planner to focus on its planning horizon and report
decision on flaws that are beyond this scope. As time advances, these
potential future flaws will eventually be part of the planning window
of the reactor and will be then resolved along with potentially new
goals.  This in turn, results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

@@

\item The token on the {\em external} timelines of this partial plan
  can then be posted which eventually (at the end of the current tick
  if these goals overlap the owner reactor's planning window) will be
  dispatched as goals, in turn triggering further deliberation.

@@

As \kcomment{shown in Section \ref{sec:arch:synch}}, it is
synchronization which drives planning; so the co-mingling between
these two processes is critical to maintaining state within a robot
especially when it is immersed in a dynamic environment. Further, it
is important that \rx maintain one $plan$ structure shared between
synchronization and \kcomment{deliberation}. This allows these two
processes to not only be interleaved for agent responsiveness, but
ensures a viable mechanism to communicate.

As \kcomment{also} noted earlier, synchronization propagates {\em
  external} observations when a new tick occurs. Earlier, we isolated
the case when a reactor has no future goal or no plan to enforce. By
doing so we were able to develop the way we augment the \eu solver to
propagate {\em external} observations to identify {\em internal}
state. This allowed us to focus on synchronization as a state
evaluation process that can be resolved within the \eu framework with
few extensions. Similarly we presented deliberation as an abstraction
without discussing the disruptive nature of synchronization and how it
impacts the planning process. \kcomment{That discussion occurs here in
  this section.}

We first analyze the case when synchronization occurs after planning
has completed (\ie the last step of deliberation results in a plan
that is considered complete for the current look-ahead). As
synchronization starts with a plan, this deliberation will not only be
a model-derived estimation uniquely based on {\em external}
observations but will also include reactor intentions specified in the
available partial plan.

Consider for example the Shopping Agent model where the timeline 
\texttt{Agent.position} is {\em external} while \texttt{Agent} is
implemented a an {\em internal} timeline. As the reactor adds a goal to
\texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable internally.

\begin{figure}[!b]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    Fig. \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for \kcomment{and tokens in
        {\color{grey}grey} the result of deliberation}. Arrows
      indicate temporal  constraints \kcomment{with $==$ indicating 'equality' and $\geq$
        indicating 'greater than equal to' in timepoints}.}
  \label{fig:shop:exec}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs the reactor of this observation which is
added to the partial plan as a fact -- starting exactly at $\tau'$ and
lasting for an unknown duration which is at least one tick. As this
new observation is integrated in the plan through synchronization, the
solver can identify the similarity between this observation and the
next token in the existing plan. Instead of initiating an exhaustive
search of all possible implications, the synchronization process
starts its execution with the assumption that these new observations
are consistent with the currently maintained plan. As a consequence,
at the end of synchronization the plan depicted in Fig.
\ref{fig:shop:exec1} is likely. This is a result of propagation of
information provided by the new observation in the existing plan while
restricting the start time of \texttt{Going} and propagating temporal
constraints, making \texttt{Go} to start {\em exactly} at $\tau'$. In
doing so, the \eu solver's focus on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}) takes advantage of the existing plan to
evaluate a solution that assumes that the observation is consistent
with what was decided by the reactor. \kcomment{In the event of
  inconsistency, the (plan) expectations do not match the reality;
  then partial plan failure within the reactor is initiated.} This is
directed not only by heuristic choices made in the solver but also
enforced by the heuristic order of the decision point resolution
choices for the {\em current state flaw} \kcomment{as} shown in
Proposition \ref{prop:csf:resolve}. As noted, the first two choices
are biased towards a conservative view with respect to the existing
plan.

@@

The \kcomment{solver's} heuristics first attempt to extend the
previous observation for this {\em internal} timeline to avoid early
commitment. If this fails \rx attempts to start the next token in the
plan to advance plan execution. Both choices attempt to resolve state
with the least amount of plan perturbation by constraining further the
start and end time of existing tokens generated by previous
deliberations steps. By doing so, in a nominal situation,
synchronization remains limited to restricting time points in the plan
and propagates their impact. Most of the computational effort of
creating new tokens in the plan and scheduling them onto timelines are
managed during deliberation steps. Consequently, most of the time
synchronization is limited to very few nodes in the \eu search space
often requiring no backtracking. Fig. \ref{fig:shop:exec} shows how
the existing plan can support the synchronization process. The
observation \texttt{Going} occurring at $\tau'$ is fully compatible
with the token already present in our plan. By updating the plan and
propagating equality constraints, the resolution of the {\em current
  state flaw} for the {\em Agent} timeline becomes obvious as a
solution, as the token \texttt{Go} from the plan presents itself.

@@

Synchronization also supports the tracking of plan execution. It does
so by specifying token parameters such as time points that are
propagated throughout the plan to identify plan validity.
%  -- as in our example -- or that it cannot be executed as it is. 
When invalid, identification occurs by synchronization failing to find
a consistent solution. For example should the \texttt{Going}
observation received be to the \texttt{hardwareStore}, this would
break the partial plan in the reactor as the robot cannot possible be
going simultaneously to two different locations; see Fig.
\ref{fig:shop:relax1}. As synchronization is not possible in such a
situation the strategy followed is to relax all decisions made by
previous plan steps. This is done by keeping all past observations in
both {\em external} and {\em internal} state variables as well as the
existing {\em internal} goals (\texttt{Own(Milk)} in this example)
received by the reactor. The resulting partial plan is shown in Fig.
\ref{fig:shop:relax2}. In this process the reactor informs the agent
that all {\em external} goals it requested for the future are no
longer valid, allowing timeline owners to divest maintaining such
goals.

\begin{figure}[!b]
  \centering
  \subfloat[\small Conflicting Observation.]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[\small Relaxed Partial Plan.]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}

>>>>

\begin{figure}[b]
  \centering
  \includegraphics[scale=2.5]{figs/token-disptch.jpeg}
  \caption{\small \kcomment{Dispatching a token occurs only when the
      request start time overlaps a window. \texttt{Goal$_1$}
      therefore gets dispatched from \texttt{Reactor$_1$} to
      \texttt{Reactor$_2$} but not \texttt{Goal$_2$} to avoid early
      commitment. Tokens from \texttt{Reactor$_1$} are 'dispatched'
      for execution to \texttt{Reactor$_2$} only when necessary while
      meeting the conditions in Definition \ref{def:dispatch}. $\tau$
      signifies the execution frontier as time moves from left to
      right in the illustration.}}
  \label{fig:reactor-lookahead}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines).]{\includegraphics[scale=2.65]{figs/AUV-agent.jpeg}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow.jpeg}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

% \label{sec:delib-exec-as}
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \subfloat[\small \kcomment{A view of synchronization and deliberation within a
  tick cycle.}]{\label{fig:sync-cycle}\includegraphics[width=0.35\columnwidth]{figs/sync-tick-cycle.jpeg}} 
  \qquad
  \subfloat[\small A \rx reactor's execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.]{\label{fig:tick-exec}\includegraphics[width=0.45\columnwidth]{figs/tick-cycle.jpeg}}
    \caption{\small \kcomment{Detailed and abstracted views of
        synchronization and deliberation within one and multiple tick
        cycles.}}
  \label{fig:sync-plan-cycle}
\end{figure}

@@

\begin{figure}[b]
  \centering \subfloat[\small When the \texttt{Sampling} reactor fails
  to synchronize, it is removed from the reactor dependency graph
  along with all its {\em internal} timelines while generating a
  \texttt{Failed} observation.]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.jpeg}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize too leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1.jpeg}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent \kcomment{from} Fig. \ref{fig:agent}.}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation.jpeg}
  \caption{\small Illustration of synchronization flaws in a
    reactor. The reactor receives new observations when they are
    produced by the owner(s) of its internal timelines. The open
    ``holes'' after the last token of each timeline represents the
    domain of all possible values for the end of this token. At every
    tick $\tau$ the reactor needs to integrate the {\em External}
    state information it received and -- based on its model
    $\mathcal{M}$ -- resolve its {\em Internal} state that will then
    be provided \kcomment{in turn} by the agent to other reactors.}
  \label{fig:synch:flaw}
\end{figure}

@@

Steps \ref{itm2} \& \ref{itm3} above are handled by the \eu solver
from Algorithm \ref{alg:europa:solve} which ensures that the {\em
  current state flaws} are selected last so as to follow the sequence
given above.

As {\em synchronization} focuses on the execution frontier it also
filters out all flaws that are outside of the current tick's scope.
If the flaw correlates to a token that is {\em necessarily} ending
before $\tau$ or starting after $\tau+1$ then the solver ignores
it. Such a strategy helps to reduce the scope of synchronization
exclusively to the current state of the reactor and in turn reduces
the search space for the {\em synchronization} solver.  To do so we
need to assume that the current state value of an {\em internal}
timeline does not depend on the future or more accurately that any
choice made during this {\em synchronization} will not lead to a
future inconsistency.  Such an assumption impacts the set of possible
domains the reactor can support while remaining
\kcomment{algorithmically} complete.

\begin{figure}[h!]
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   equals(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11 }
\end{verbatim}
\label{fig:nddl-sync}
\caption{\small \nd rule from the Shopping Agent example with impact
  on a future state.}
\end{figure}

Take for example this rule from our \texttt{Shopping} model in
Fig. \ref{fig:nddl-sync}. While this model is perfectly acceptable for
planning it becomes problematic when execution is taken into account
especially since it puts a strong emphasis between an evolving plan
and an unknown future; \texttt{Going} to a location and the {\em
  expected} future outcome of ending at this location, ties it
strongly to the constraint in line $5$ above. For instance, assume
that the \texttt{Agent.location} is {\em external} to the reactor
which receives the observation \texttt{Going(Home,SuperMarket)} which
synchronization has resolved by producing the observation
\texttt{Go(Home, SuperMarket)} in the \texttt{Agent} timeline. This
would imply that the only possible outcome on the
\texttt{Agent.location} timeline, is to be \texttt{At(SuperMarket)} in
the directly foreseeable future.

In the real world the fact that the robot is \texttt{Going} to a
certain location does not necessarily imply that it will succeed in
ending up there. Due to an exogenous event, for example, the road to
the \texttt{SuperMarket} could be closed, which could result in an
alternate observation that the robot is \texttt{At(7th st. \& Main)}
contradicting the existing plan. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes the
contradiction of the \texttt{At} predicate, the rule that generated
this conflict is connected to past observations \texttt{Going} and
\texttt{Go}. These cannot be retracted easily without a potential
impact on computational cost.

@@

One could therefore alter this part of the model to give the reactor
more flexibility by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ to ensure that the last \texttt{Going} ends up in the
location the agent actually wants the robot to be at. This first level
of relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first token fail to reach the
objective, as shown below:

\begin{figure}[tbp]
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
\caption{\small Model fragment to enforce a \texttt{Going} end condition}
\label{lab:model:shopping}
\end{figure}

As the example in Fig. \ref{lab:model:shopping} shows, synchronization
involves a deliberation process. The risk is such that the search for
a solution is not a simple sequence of decisions where the resolution
can be easily identified by the solver as any other option would lead
immediately to plan inconsistency. Since such a process is required by
the \rx at \emph{every} tick for \emph{every} reactor the model needs
to be designed carefully.

Synchronization is the only critical task of a reactor. Such a process
is used in order for the agent to identify its current state and
maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model \kcomment{being} unable to find a
consistent plan -- not only indicates that a reactor is unable to
explain the state of the world, but more importantly it may jeopardize
other reactors that rely on its {\em internal} state. For this reason
a failure to synchronize immediately, results in the \rx agent killing
a reactor and notifying any reactor that depends on its {\em internal}
state variables; it does so by posting the special observation
\textsf{Failed}. This special state is acceptable for all the {\em
  internal} and {\em external} timelines of \kcomment{an} \eu reactor.
By so doing we allow for graceful degradation where other reactors can
readapt their plans on \kcomment{receiving} this observation.

Finally, at the engine level careful choices \kcomment{are} made in
both the focus of the search -- limited only to tokens that can
potentially overlap the current tick -- and heuristics where the {\em
  current state flaw} is considered last during the search
process. The later limits the amount of backtracking for a solution --
or at least how deep one needs to backtrack in a decision tree.
Similarly the sequence in which we evaluate the possible resolution
for such a flaw -- \kcomment{described} in Proposition
\ref{prop:csf:resolve} -- \kcomment{is} selected according to
\kcomment{the assumption}, that any external observation that is
compatible with an existing token in the current plan reflects the
execution of the plan itself. In other words this means that our
solver always prefers to use tokens \emph{already existing} in the
plan as a way to identify current internal state. By doing so
synchronization exploits the structure of the plan produced during
previous deliberation steps.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window.jpeg}
  \caption{\small The planning window of the deliberation solver and
    its relation to the {\em dispatch window}. The dark grey tokens
    with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$. All the other
    tokens ({\color{green} green}) along with the goals received from
    the agent ({\color{blue} blue}) which can overlap this window, are
    evaluated.}
  \label{fig:plan:window}
\end{figure}

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch window} from
Definition \ref{def:dispatch} which filters out, at the agent level,
those goals that are before the reactor execution latency or too far
into the future.  The planning window refines this filtering further,
as tokens can be excluded dynamically from deliberation as their
insertion in the plan and scheduling with other generated tokens can
alter their temporal scope to be outside of the planning window.  This
allows the planner to focus on its planning horizon and not report
decisions on flaws that are beyond this scope. As time advances, these
potential future flaws will eventually be part of the planning window
of the reactor and will be then resolved along with potentially new
goals.  This in turn, results in an apparent \emph{continuous
  planning} \kcomment{\cite{myers99}} of the reactor. When there are
no more flaws present in the current planning scope a partial plan
solution is found. This has two effects:

@@

\item The token on the {\em external} timelines of this partial plan
  can then be posted. \kcomment{Eventually, at the end of the current
    tick if this goal overlaps the owner reactor's planning window, it
    will be dispatched, resulting in further deliberation.}

@@

As \kcomment{shown in Section \ref{sec:arch:synch}}, it is
synchronization which drives planning; so the co-mingling between
these two processes is critical to maintaining state within a robot.
Further, it is important that \rx maintain one $plan$ structure shared
between synchronization and \kcomment{deliberation}. This allows these
two processes to not only be interleaved for agent responsiveness, but
ensures a viable mechanism to communicate.

As \kcomment{also} noted earlier, synchronization propagates {\em
  external} observations when a new tick occurs. Earlier, we isolated
the case when a reactor has no future goal or no plan to enforce. By
doing so we were able to develop the way we augment the \eu solver to
propagate {\em external} observations to identify {\em internal}
state. This \kcomment{ensured} focus on synchronization as a state
evaluation process that can be resolved within the \eu framework with
few extensions. Similarly deliberation was presented as an abstraction
without discussing the disruptive nature of synchronization and how it
impacts the planning process. \kcomment{That discussion occurs here in
  this section.}

We first analyze the case when synchronization occurs after planning
has completed, \ie the last step of deliberation results in a plan
that is considered complete for the current look-ahead. As
synchronization starts with a plan, this deliberation will not only be
a model-derived estimation uniquely based on {\em external}
observations but will also include reactor intentions specified in the
partial plan.

Consider for example the Shopping Agent model \kcomment{from Section
  \ref{sec:europa:modeling}}, where the timeline
\texttt{Agent.position} is {\em external} while \texttt{Agent} is
implemented as an {\em internal} timeline. As the reactor adds a goal
to \texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable internally.

\begin{figure}[!b]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0.jpeg}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    Fig. \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1.jpeg}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for \kcomment{and tokens in
        {\color{grey}grey} the result of deliberation}. Arrows
      indicate temporal  constraints \kcomment{with $==$ indicating 'equality' and $\geq$
        indicating 'greater than equal to' in timepoints}.}
  \label{fig:shop:exec}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs the reactor \kcomment{that owns the
  \texttt{Agent} timeline,} of this observation which is added to
\kcomment{its} partial plan as a fact -- starting exactly at $\tau'$
and lasting for an unknown duration which is at least one tick. As
this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
observation and the next token in the existing plan. Instead of
initiating an exhaustive search of all possible implications, the
synchronization process starts its execution with the assumption that
these new observations are consistent with the currently maintained
plan. As a consequence, at the end of synchronization the plan
depicted in Fig.  \ref{fig:shop:exec1} is likely. This is a result of
propagation of information provided by the new observation in the
existing plan while restricting the start time of \texttt{Going} and
propagating temporal constraints, making token \texttt{Go} to start
{\em exactly} at $\tau'$. In doing so, the \eu solver's focus on the
execution frontier ($\tau'$ in Fig. \ref{fig:shop:exec1}) takes
advantage of the existing plan to evaluate a solution that assumes
that the observation is consistent with what was decided by the
reactor. \kcomment{In the event of inconsistency, plan expectations do
  not match reality; then partial plan failure within the reactor is
  initiated.} This is directed not only by heuristic choices made in
the solver but also enforced by the heuristic order of the decision
point resolution choices for the {\em current state flaw}
\kcomment{as} shown in Proposition \ref{prop:csf:resolve}. As noted,
the first two choices are biased towards a conservative view with
respect to the existing plan.

@@

The \kcomment{solver's} heuristics first attempt to extend the
previous observation for this {\em internal} timeline to avoid early
commitment. If this fails \rx attempts to start the next token in the
plan to advance plan execution. Both choices attempt to resolve state
with the least amount of plan perturbation by further constraining the
start and end time of existing tokens generated by previous
deliberations steps. By doing so, in a nominal situation,
synchronization remains limited to restricting time points in the plan
and \kcomment{in turn} propagates their impact. Most of the
computational effort of creating new tokens in the plan and scheduling
them onto timelines are managed during deliberation
steps. Consequently, most of the time synchronization is limited to
very few nodes in the \eu search space often requiring no
backtracking. Fig. \ref{fig:shop:exec} shows how the existing plan can
support the synchronization process. The observation \texttt{Going}
occurring at $\tau'$ is fully compatible with the token already
present in our plan. By updating the plan and propagating equality
constraints, the resolution of the {\em current state flaw} for the
{\em Agent} timeline becomes obvious as a solution, as the token
\texttt{Go} from the plan presents itself.

@@

Synchronization also supports the tracking of plan execution. It does
so by specifying token parameters such as time points that are
propagated throughout the plan to identify plan validity.
%  -- as in our example -- or that it cannot be executed as it is. 
When invalid, identification occurs by synchronization failing to find
a consistent solution. For instance, should the \texttt{Going}
observation received be to the \texttt{hardwareStore}, this would
break the partial plan in the reactor as the robot cannot possible be
going simultaneously to two different locations; see Fig.
\ref{fig:shop:relax1}. As synchronization is not possible in such a
situation the strategy followed is to relax all decisions made by
previous plan steps. This is done by keeping all past observations in
both {\em external} and {\em internal} state variables as well as the
existing {\em internal} goals (\texttt{Own(Milk)} in this example)
received by the reactor. The resulting partial plan is shown in Fig.
\ref{fig:shop:relax2}. In this process the reactor informs the agent
that all {\em external} goals it requested for the future are no
longer valid, allowing timeline owners to divest maintaining such
goals.

\begin{figure}[!b]
  \centering
  \subfloat[\small Conflicting Observation.]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1.jpeg}}\\
  \subfloat[\small Relaxed Partial Plan.]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2.jpeg}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}
